# Regression analysis

*Second week, I did the data wrangling exercise and build small regression analysis on given data.*

First, I imported the libraries and got data from URL to data frame

```{r}

# Access the dplyr and ggplot2 libraries
library(dplyr)
library(ggplot2)
library(GGally)

# Read data from URL

my_data <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/learning2014.txt", sep=",", header=TRUE)



dim(my_data)

str(my_data)

```
The result shows that the data has 166 rows and 7 variables. Seems as intended.

Next I did summary visualizations just to get understanding of the data

```{r}
# create advanced plot matrix with ggpairs()
p <- ggpairs(my_data, mapping = aes(col=gender), lower = list(combo = wrap("facethist", bins = 20)))

# draw the plot
p
```

Deep, strat and surf are close to normally distributed.

Age seems far from normally distributed, which is easy to understand because most of the students are young.

Differences between genders seem small.

Highest correlation with points is in attitude (0.437), stra (0.146) and surf (-1.44)

For regression model, I chose attitude, stra and age

```{r}

# create a regression model with multiple explanatory variables
my_model <- lm(points ~ attitude + stra + age, data = my_data)

# print out a summary of the model
summary(my_model)
```
Seems like only variable having statistical significance is attitude (p<0.001)

Let's try simple regression with only attitude variable

```{r}
# fit a linear model
my_model2 <- lm(points ~ attitude, data = my_data)

# print out a summary of the model
summary(my_model2)

```
Model is statistically significant (p<0.001)

Alfa has a value 11.6 and increase of one in attitude seems to increase points by 3.5

Multiple R-squared is 0.19 which means that attitude explains roughly one fifth of the changes in points

The main conclusion is that differences in attitude explain quite well the difference in course grade (points)

The last thing is to draw few diagnostic plots:

```{r}
# draw diagnostic plots using the plot() function. Choose the plots 1, 2 and 5 (Residuals vs Fitted values, Normal QQ-plot and Residuals vs Leverage)
par(mfrow = c(2,2))
plot(my_model2, which = c(1, 2, 5))
```

Most of the residuals seem small

Observations follow the regression line nicely

There is no single observation that would draw the regression line heavily in some direction

Everything seems fine
