---
title: "chapter4"
author: "Petri Heinonen"
date: "2/24/2017"
output: html_document
---


# Dimensionality reduction 

This is the fifth chapter including dimensionality reduction techniques.

First, I read the data from URL and explored basic structure and dimensions
```{r}

# Read data from URL

human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt", sep=",", header=TRUE)

str(human)
dim(human)

```

Data has 155 observations and 8 variables. All are either numeric or integers.

Let's visualize the data and check some correlations.
```{r}

# Access GGally
library(GGally)
library(corrplot)

# visualize the 'human' variables
ggpairs(human)

# compute the correlation matrix and visualize it with corrplot
cor(human) %>% corrplot()
```

Expected education is somewhat normally distributed. All the others have differ quite a lot from normal distribution.

Maternal mortality and life expectancy have strong negative correlation which is easy to understand. Education expectancy and life expectancy have highest positive correlation. It would be interesting to study whether it's causality.

Next some PCA.

```{r}

# perform principal component analysis
pca_human <- prcomp(human)

# draw a biplot of the principal component representation and the original variables
biplot(pca_human, choices = 1:2, cex = c(0.8, 1), col = c("grey40", "deeppink2"))
```

It's hard to say anything about the non-standardized. Needs some scaling.

```{r}

# standardize the variables
human_std <- scale(human)

# perform principal component analysis
pca_human <- prcomp(human_std)

# draw a biplot of the principal component representation and the original variables
biplot(pca_human, choices = 1:2, cex = c(0.8, 1), col = c("grey40", "deeppink2"))
```

Seems that labor ratio between sexes and the female proportion in parliament draws heavily in the same direction as PC2. Western countries and African countries have high PC2. Islamic countries have low PC2. PC2 is probably a dimension that summaries variables that show gender equality.

Education and life expectancy variables correlate positively with PC1. Birth rate and maternal mortality correlate negatively. PC1 seems to combine variables that show general development level of the country.